### 2018-01-12 epss_marmo.tex

%% 2/18/2016
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% AGUJournalSample.tex: this sample file is for articles formatted with LaTeX
%
% This sample file includes commands and instructions
% given in the order necessary to produce a final output that will
% satisfy AGU requirements.
%
% PLEASE DO NOT USE YOUR OWN MACROS
% DO NOT USE \newcommand, \renewcommand, or \def.
%
% FOR FIGURES, DO NOT USE \psfrag or \subfigure.
% DO NOT USE \psfrag or \subfigure commands.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% All questions should be e-mailed to latex@agu.org.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Step 1: Set the \documentclass
%
% There are two options for article format:
%
% 1) PLEASE USE THE DRAFT OPTION TO SUBMIT YOUR PAPERS.
% The draft option produces double spaced output.
% 
% 2) numberline will give you line numbers.

% Tip:
%  To add line numbers to lines in equations:
%  \begin{linenomath*}
%  \begin{equation}
%  \end{equation}
%  \end{linenomath*}

%% To submit your paper:
\documentclass[linenumbers,draft]{agujournal}

% Now, type in the journal name: \journalname{<Journal Name>}
% ie,
\journalname{Earth and Space Science}

%% Choose from this list of Journals:
%
% JGR-Atmospheres
% JGR-Biogeosciences
% JGR-Earth Surface
% JGR-Oceans
% JGR-Planets
% JGR-Solid Earth
% JGR-Space Physics
% Global Biochemical Cycles
% Geophysical Research Letters
% Paleoceanography
% Radio Science
% Reviews of Geophysics
% Tectonics
% Space Weather
% Water Resource Research
% Geochemistry, Geophysics, Geosystems
% Journal of Advances in Modeling Earth Systems (JAMES)
% Earth's Future
% Earth and Space Science

%% ------------------------------------------------------------------------ %%
%
%  ENTER Title Page commands:
%
%% ------------------------------------------------------------------------ %%

% (A title should be specific, informative, and brief. Use
% abbreviations only if they are defined in the abstract. Titles that
% start with general terms then specific results are optimized in
% searches)

% Example: \title{This is a test title}

% (List authors by first name or initial followed by last name and
% separated by commas. Use \affil{} to number affiliations, and
% \thanks{} for author notes.  
% Additional author notes should be indicated with \thanks{} (for
% example, for current addresses).

% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
% Author\affil{3,4}\thanks{Also funded by Monsanto.}}

% (include name and email addresses of the corresponding author.  More
% than one corresponding author is allowed in this LaTeX file and for
% publication; but only one corresponding author is allowed in our
% editorial system.)

%% Corresponding Author:
% Corresponding author mailing address and e-mail address:

% Example: \correspondingauthor{First and Last Name}{email@address.edu}

% Authors are individuals who have significantly contributed to the
% research and preparation of the article. Group authors are allowed, if
% each author in the group is separately identified in an appendix.)

% \affiliation{1}{First Affiliation}
% \affiliation{2}{Second Affiliation}
% \affiliation{3}{Third Affiliation}
% \affiliation{4}{Fourth Affiliation}

%% Keypoints, final entry on title page.
% Example: 
% \begin{keypoints}
% \item	List up to three key points (at least one is required)
% \item	Key Points summarize the main points and conclusions of the article
% \item	Each must be 100 characters or less with no special
% characters or punctuation 
% \end{keypoints}

%% \begin{abstract} begins second page

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Track Changes:
% To add words, \added{<word added>}
% To delete words, \deleted{<word deleted>}
% To replace words, \replace{<word to be replaced>}{<replacement word>}
% To explain why change was made: \explain{<explanation>}

% At the end of the document, use \listofchanges, which will list the
% changes and the page and line number where the change was made.

% When final version, \listofchanges will not produce anything,
% \added{} word will be printed, \deleted{} will take away the word,
% \replaced{}{} will print only the 2nd argument.
% \explain will not print anything.

% Optional argument:
% You can also add additional information to be printed with the list
% of changes, to indicate the initials of the person changing the text,
% and the time and/or date of the change, or any other comment by using
% the optional [] argument:
% \added[AH, 3:30pm, Feb 18, 2016]{added term}
% will yield 
% [AH, 3:30pm, Feb 18, 2016] added term on page...
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%% ------------------------------------------------------------------------ %%
%
%  TITLE
%
%% ------------------------------------------------------------------------ %%

\title{FITS format for planetary surfaces: definitions, applications and best practices.}

%% ------------------------------------------------------------------------ %%
%
%  AUTHORS AND AFFILIATIONS
%
%% ------------------------------------------------------------------------ %%

 \authors{C. Marmo\affil{1},
 T. M. Hare\affil{2}, ...}

\affiliation{1}{GEOPS, Univ. Paris-Sud, CNRS, Univ. Paris-Saclay, Orsay, France}
\affiliation{2}{U. S. Geological Survey, Astrogeology Science Center, Flagstaff, AZ, USA}
%\affiliation{3}{}
%\affiliation{4}{}
%\affiliation{5}{}

%% Corresponding Author
%(include name and email addresses of the corresponding author.  More
%than one corresponding author is allowed in this Word file and for
%publication; but only one corresponding author is allowed in our
%editorial system.)

\correspondingauthor{C. Marmo}{chiara.marmo@u-psud.fr}

%  List up to three key points (at least one is required)
%  Key Points summarize the main points and conclusions of the article
%  Each must be 100 characters or less with no special characters or punctuation

\begin{keypoints}
\item Inter-operable data formats 
\item Planetary data standards
\item Data processing and visualization
\end{keypoints}

%% ------------------------------------------------------------------------ %%
%
%  ABSTRACT
%
%% ------------------------------------------------------------------------ %%

\begin{abstract}

\end{abstract}

%% ------------------------------------------------------------------------ %%
%
%  TEXT
%
%% ------------------------------------------------------------------------ %%

%%% Suggested section heads:
% \section{Introduction}
% 
% The main text should start with an introduction. Except for short
% manuscripts (such as comments and replies), the text should be divided
% into sections, each with its own heading.

% Headings should be sentence fragments and do not begin with a
% lowercase letter or number. Examples of good headings are:

% \section{Materials and Methods}
% Here is text on Materials and Methods.

% \subsection{A descriptive heading about methods}
% More about Methods.
% 
% \section{Data} (Or section title might be a descriptive heading about data)
% 
% \section{Results} (Or section title might be a descriptive heading about the
% results)
% 
% \section{Conclusions}

\section{Introduction}
\label{sec:intro}

%% Enter Figures and Tables near as possible to where they are first mentioned:
%
% DO NOT USE \psfrag or \subfigure commands.
%
% Figure captions go below the figure.
% Table titles go above tables;  other caption information
%  should be placed in last line of the table, using
% \multicolumn2l{$^a$ This is a table note.}
%
%----------------
% EXAMPLE FIGURE
%
% \begin{figure}
% \end{figure}

\begin{figure}[ht!]
% \centering
% when using pdflatex, use pdf file:
% \includegraphics[width=20pc]{figsamp.pdf}
%
% when using dvips, use .eps file:
% \includegraphics[width=20pc]{figsamp.eps}
%
% If you don't specify the file extension, 
% \includegraphics will insert the right file:
\centerline{\includegraphics[height=1.5in]{figsamp}}
\caption{Short caption}
\label{figone}
\end{figure}

\begin{figure}[ht!]
\centerline{\includegraphics[angle=45,height=2in]{figsamp}}
\caption{The figure caption should begin with an overall descriptive
statement of the figure followed by additional text. They should be
immediately after each figure.  Figure parts are indicated with
lower-case letters ({\bf a, b, c}\ldots).  For initial submission, please place
both the figures and captions in the text near where they are cited.}
\label{fig:mean_and_slope}
\end{figure}

\begin{table}
\caption{Start this caption with a short description of your
table.
Large tables
especially presenting rich data should be presented as separate excel
or .cvs files, not as part of the main text.}
== Table Here ==
\end{table}

%
% ---------------
% EXAMPLE TABLE
%
\begin{table}
\caption{Time of the Transition Between Phase 1 and Phase 2$^{a}$}
\centering
\begin{tabular}{l c}
\hline
 Run  & Time (min)  \\
\hline
  $l1$  & 260   \\
  $l2$  & 300   \\
  $l3$  & 340   \\
  $h1$  & 270   \\
  $h2$  & 250   \\
  $h3$  & 380   \\
  $r1$  & 370   \\
  $r2$  & 390   \\
\hline
\multicolumn{2}{l}{$^{a}$Table note text here.}
\end{tabular}
\end{table}

% See below for how to make sideways figures or tables.

% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
%
 \begin{sidewaysfigure}
 \includegraphics[width=20pc]{figsamp}
 \caption{Caption here}
 \label{newfig}
 \end{sidewaysfigure}
%
 \begin{sidewaystable}
 \caption{Caption here}
\label{tab:signif_gap_clos}
 \begin{tabular}{ccc}
one&two&three\\
four&five&six
 \end{tabular}
 \end{sidewaystable}

\clearpage

\section{FITS format in Planetary Surface investigation}
\label{sec:fitspss}

\subsection{FITS format}

\subsubsection{Planetary Surface proposed convention}

\subsubsection{FITS to PDS3 and PDS4 dictionaries}

\subsubsection{Best Practices and Examples}

\section{Inter-operable software developments}
\label{sec:softdev}

\section{Discussion and Perspectives}
\label{sec:disc}

%% ------------------------------------------------------------------------ %%
\section*{Citations}

% Please use ONLY \citet and \citep for reference citations.
% DO NOT use other cite commands (e.g., \cite, \citeyear, \nocite, \citealp, etc.).

\subsection*{Cites made with \tt\string\citet\string{\string}}
...as shown by \citet{Boug10}, \citet{Buiz07}, \citet{Fra10},
\citet{Ghel00}, and \citet{Leit74}.

\subsection*{Cites made with \tt\string\citep\string{\string}}
...as shown by \citep{Boug10}, \citep{Buiz07}, \citep{Fra10},
\citep{Ghel00, Leit74}.

...has been shown \citep [e.g.,][]{Boug10,Buiz07,Fra10}.

%%% End of body of article

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional Appendix goes here
%
% \appendix resets counters and redefines section heads
% but doesn't print anything.
% After typing \appendix
%
%\section{Here Is Appendix Title}
% will show
% A: Here Is Appendix Title
%
\appendix
\section{Here is a sample appendix}
This is an Appendix section.

\subsection{subsection}
This is an Appendix subsection.

\subsubsection{subsubsection}
This is an Appendix subsubsection.
\begin{linenomath*}
\begin{equation}asdf\end{equation}
\end{linenomath*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Optional Glossary, Notation or Acronym section goes here:
%
%%%%%%%%%%%%%%
% Glossary is only allowed in Reviews of Geophysics
\begin{glossary}
\term{Term}
 Term Definition here
\term{Term}
 Term Definition here
\term{Term}
 Term Definition here
\end{glossary}

%
%%%%%%%%%%%%%%
% Acronyms
 \begin{acronyms}
 \acro{Acronym}
 Definition here
 \acro{EMOS}
 Ensemble model output statistics 
 \acro{ECMWF}
 Centre for Medium-Range Weather Forecasts
 \end{acronyms}

%
%%%%%%%%%%%%%%
% Notation 
 \begin{notation}
 \notation{$a+b$} Notation Definition here
 \notation{$e=mc^2$} 
 Equation in German-born physicist Albert Einstein's theory of special
relativity that showed that the increased relativistic mass ($m$) of a
body comes from the energy of motion of the body—that is, its kinetic
energy ($E$)—divided by the speed of light squared ($c^2$).
 \end{notation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  ACKNOWLEDGMENTS

\acknowledgments
This work benefits from support of VESPA/Europlanet.
The Europlanet 2020 Research Infrastructure is funded by the European Union under the
Horizon 2020 research and innovation program, grant agreement N.654208.

%%  REFERENCE LIST AND TEXT CITATIONS
%
% Either type in your references using
%
% \begin{thebibliography}{}
% \bibitem{}
% Text
% \end{thebibliography}
%
% Or, to use BibTeX:
%
% Follow these steps
%
% 1. Type in \bibliography{<name of your .bib file>} 
%    Run LaTeX on your LaTeX file.
%
% 2. Run BiBTeX on your LaTeX file.
%
% 3. Open the new .bbl file containing the reference list and
%   copy all the contents into your LaTeX file here.
%
% 4. Run LaTeX on your new file which will produce the citations.
%

\bibliography{epss_marmo}

% AGU does not want a .bib or a .bbl file. Please copy in the contents of your .bbl file here.

% % \begin{thebibliography}{}

% % \bibitem[{\textit{Bell and Munoz}}(2008)]{Boug10} Bell, A.~H., and
% % Munoz, D.~P.  (2008). Activity in the superior colliculus reflects
% % dynamic interactions between voluntary and involuntary influences on
% % orienting behaviour. \textit{Eur. J. Neurosci.} 28, 1654--1660.

% % \bibitem[{\textit{Corbetta et~al.}}(1991)]{Buiz07} Corbetta, M.,
% % Miezin, F.~M., Dobmeyer, S., Shulman, G.~L., and Petersen, S.~E.
% % (1991). Selective and divided attention during visual discriminations
% % of shape, color, and speed: functional anatomy by positron emission
% % tomography. \textit{J.~Neurosci.} 11, 2383--2402.

% % \bibitem[{\textit{Borra et~al.}}(2014)]{Buiz98} Borra, E., Gerbella,
% % M., Rozzi, S., Tonelli, S., and Luppino, G.  (2014). Projections to
% % the superior colliculus from inferior parietal, ventral premotor, and
% % ventrolateral prefrontal areas involved in controlling goal-directed
% % hand actions in the macaque. \textit{Cereb. Cortex} 24, 1054--1065.

% % \bibitem[{\textit{Dorris et~al.}}(1997)]{Fra10}
% %  Dorris, M.~C.,
% % Par\'{e}, M., and Munoz, D.~P.  (1997). Neuronal activity in monkey
% % superior colliculus related to the initiation of saccadic eye
% % movements. \textit{J.~Neurosci.} 17, 8566--8579.

% % \bibitem[{\textit{Elsabbagh et~al.}}(2009)]{Ghel00} 
% % Elsabbagh, M.,
% % Volein, A., Holmboe, K., Tucker, L., Csibra, G., Baron-Cohen, S.,
% % et~al.  (2009). Visual orienting in the early broader autism
% % phenotype: disengagement. \textit{J.~Child Psychol. Psychiatry} 50,
% % 637--642.

% % \bibitem[{\textit{Fortin et~al.}}(1999)]{Gneit05} Fortin, S., Chabli,
% % A., Dumont, I., Shumikhina, S., Itaya, S.~K., and Molotchnikoff, S.
% % (1999). Maturation of visual receptive field properties in the rat
% % superior colliculus. \textit{bibain Res. Dev. Brain Res.} 1112,
% % 55--64.

% % \bibitem[{\textit{Felsen and Mainen}}(2008)]{Gneit07b} Felsen, G., and
% % Mainen, Z.~F.  (2008). Neural substrates of sensory-guided locomotor
% % decisions in the rat superior colliculus. \textit{Neuron} 60,
% % 137--148.

% % \bibitem[{\textit{Gattass and Desimone}}(1996)]{Haid14} Gattass, R.,
% % and Desimone, R.  (1996). Responses of cells in the superior
% % colliculus during performance of a spatial attention task in the
% % macaque. \textit{Rev. Bras. Biol.} 56, 257--279.

% % \bibitem[{\textit{Goldberg and Wurtz}}(1972)]{Hami00} Goldberg, M.~E.,
% % and Wurtz, R.~H.  (1972). Activity of superior colliculus in behaving
% % monkey. II. Effect of attention on neuronal responses.
% % \textit{J.~Neurophysiol.} 35, \hbox{560--574}.

% % \bibitem[{\textit{Krauzlis}}(2003)]{Kendall} Krauzlis, R.~J.  (2003).
% % Neuronal activity in the rostral superior colliculus related to the
% % initiation of pursuit and saccadic eye movements.
% % \textit{J.~Neurosci.} 23, 4333--4344.

% % \bibitem[{\textit{Heesy}}(2009)]{Leit74} 
% % Heesy, C.~P.  (2009). Seeing
% % in stereo: the ecology and evolution of primate binocular vision and
% % stereopsis. \textit{Evol. Anthropol.} 18, 21--35.

% % \bibitem[{\textit{Hilbig et~al.}}(2000)]{Mann45} Hilbig, H., Bidmon,
% % H.~J., Ettrich, P., and M\"{u}ller, A.  (2000). Projection neurons in
% % the superficial layers of the superior colliculus in the rat: a
% % topographic and quantitative morphometric analysis.
% % \textit{Neuroscience} 96, 109--119.

% % \bibitem[{\textit{Ignashchenkova et~al.}}(2004)]{Math76}
% % Ignashchenkova, A., Dicke, P.~W., Haarmeier, T., and Their, P.
% % (2004). Neuron-specific contribution of the superior colliculus to
% % overt and covert shifts of attention. \textit{Nat. Neurosci.} 7,
% % 56--64.

% \bibitem[{\textit{Krauzlis et~al.}}(2013)]{Palm00} Krauzlis, R.~J.,
% Lovejoy, L.~P., and Z\'{e}non, A.  (2013). Superior colliculus and
% visual spatial attention. \textit{Annu. Rev. Neurosci.} 36, 165--182.

% \bibitem[{\textit{Kustov and Robinson}}(1996)]{Papp09} Kustov, A.~A.,
% and Robinson, D.~L.  (1996). Shared neural control of attentional
% shifts and eye movements. \textit{Nature} 384, 74--77.

% \bibitem[{\textit{Landry and Bryson}}(2004)]{Park08} Landry, R., and
% Bryson, S.~E.  (2004). Impaired disengagement of attention in young
% children with autism. \textit{J.~Child Psychol. Psychiatry} 45,
% 1115--1122.

% \bibitem[{\textit{Kobayashi et~al.}}(2003)]{R2013} Kobayashi, T.,
% Tran, A.~H., Nishijo, H., Ono, T., and Matsumoto, G.  (2003).
% Contribution of hippocampal place cell activity to learning and
% formation of goal-directed navigation in rats. \textit{Neuroscience}
% 117, 1025--1035.

% \bibitem[{\textit{McHaffie et~al.}}(2005)]{Raft05} McHaffie, J.~G.,
% Stanford, T.~R., Stein, B.~E., Coizet, V., and Redgrave, P.  (2005).
% Subcortical loops through the basal ganglia. \textit{Trends Neurosci.}
% 28, 401--407.

% \bibitem[{\textit{McPeek and Keller}}(2004)]{Rich13} McPeek, R.~M.,
% and Keller, E.~L.  (2004). Deficits in saccade target selection after
% inactivation of superior colliculus. \textit{Nat. Neurosci.} 7,
% 757--763.

% \bibitem[{\textit{M\"{u}ller et~al.}}(2005)]{Scheu14} M\"{u}ller,
% J.~R., Philiastides, M.~G., and Newsome, W.~T.  (2005).
% Microstimulation of the superior colliculus focuses attention without
% moving the eyes. \textit{Proc. Natl. Acad. Sci. U.S.A.} 102, 524--529.

% \bibitem[{\textit{Munoz and Istvan}}(1998)]{ScheueBue13} Munoz, D.~P.,
% and Istvan, P.~J.  (1998). Lateral inhibitory interactions in
% the intermediate layers of the monkey superior colliculus.
% \textit{J.~Neurophysiol.} 79, 1193--1209.

% \end{thebibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Track Changes:
% To add words, \added{<word added>}
% To delete words, \deleted{<word deleted>}
% To replace words, \replace{<word to be replaced>}{<replacement word>}

% At the end of the document, use \listofchanges, which will list the
% changes and the page and line number where the change was made.

% When final version, \listofchanges will not produce anything,
% \added{} word will be printed, \deleted{} will take away the word,
% \replaced{}{} will print only the 2nd argument.

%%%
\listofchanges
%%%

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Supporting Information
%% (Optional) See AGUSuppInfoSamp.tex/pdf for requirements 
%% for Supporting Information.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

More Information and Advice:

%% ------------------------------------------------------------------------ %%
%
%  SECTION HEADS
%
%% ------------------------------------------------------------------------ %%

% Capitalize the first letter of each word (except for
% prepositions, conjunctions, and articles that are
% three or fewer letters).

% AGU follows standard outline style; therefore, there cannot be a section 1 without
% a section 2, or a section 2.3.1 without a section 2.3.2.
% Please make sure your section numbers are balanced.
% ---------------
% Level 1 head
%
% Use the \section{} command to identify level 1 heads;
% type the appropriate head wording between the curly
% brackets, as shown below.
%
%An example:
%\section{Level 1 Head: Introduction}
%
% ---------------
% Level 2 head
%
% Use the \subsection{} command to identify level 2 heads.
%An example:
%\subsection{Level 2 Head}
%
% ---------------
% Level 3 head
%
% Use the \subsubsection{} command to identify level 3 heads
%An example:
%\subsubsection{Level 3 Head}
%
%---------------
% Level 4 head
%
% Use the \subsubsubsection{} command to identify level 3 heads
% An example:
%\subsubsubsection{Level 4 Head} An example.
%
%% ------------------------------------------------------------------------ %%
%
%  IN-TEXT LISTS
%
%% ------------------------------------------------------------------------ %%
%
% Do not use bulleted lists; enumerated lists are okay.
% \begin{enumerate}
% \item
% \item
% \item
% \end{enumerate}
%
%% ------------------------------------------------------------------------ %%
%
%  EQUATIONS
%
%% ------------------------------------------------------------------------ %%

% Single-line equations are centered.
% Equation arrays will appear left-aligned.

Math coded inside display math mode \[ ...\]
 will not be numbered, e.g.,:
 \[ x^2=y^2 + z^2\]

 Math coded inside \begin{equation} and \end{equation} will
 be automatically numbered, e.g.,:
 \begin{equation}
 x^2=y^2 + z^2
 \end{equation}

% To create multiline equations, use the
% \begin{eqnarray} and \end{eqnarray} environment
% as demonstrated below.
\begin{eqnarray}
  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
        && + (y - y_{0}) \sin \Theta  \nonumber \\
  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
        && + (y - y_{0}) \cos \Theta.
\end{eqnarray}

%If you don't want an equation number, use the star form:
%\begin{eqnarray*}...\end{eqnarray*}

% Break each line at a sign of operation
% (+, -, etc.) if possible, with the sign of operation
% on the new line.

% Indent second and subsequent lines to align with
% the first character following the equal sign on the
% first line.

% Use an \hspace{} command to insert horizontal space
% into your equation if necessary. Place an appropriate
% unit of measure between the curly braces, e.g.
% \hspace{1in}; you may have to experiment to achieve
% the correct amount of space.

%% ------------------------------------------------------------------------ %%
%
%  EQUATION NUMBERING: COUNTER
%
%% ------------------------------------------------------------------------ %%

% You may change equation numbering by resetting
% the equation counter or by explicitly numbering
% an equation.

% To explicitly number an equation, type \eqnum{}
% (with the desired number between the brackets)
% after the \begin{equation} or \begin{eqnarray}
% command.  The \eqnum{} command will affect only
% the equation it appears with; LaTeX will number
% any equations appearing later in the manuscript
% according to the equation counter.
%

% If you have a multiline equation that needs only
% one equation number, use a \nonumber command in
% front of the double backslashes (\\) as shown in
% the multiline equation above.

% If you are using line numbers, remember to surround
% equations with \begin{linenomath*}...\end{linenomath*}

%  To add line numbers to lines in equations:
%  \begin{linenomath*}
%  \begin{equation}
%  \end{equation}
%  \end{linenomath*}