### 2018-05-08 2018EA000388_Answers_1Reviewer.txt
Working on ESS first internal review.

2018EA000388 Answers to the first Reviewer

l. 43
Rephrased: see the new version of the manuscript

l. 57
We agree with the reviewer: the paragraph has been deleted

l. 68
Rephrased: see the new version of the manuscript

l. 68-69
IPDA informally discussed about proposing recommended data formats for archive.
As an IPDA official position is missing we cannot argue about FITS in that context.

l. 83-84

l. 102
GDAL lack of metadata description for FITS format: added in the manuscript

Table 1: note b

Table 2 : header
Mercury must be there... sorry for the oversight.
TNO and KBO are Asteroids: this is now clarified in the note.