### 2018-05-09 2018EA000388_Answers_1Reviewer.txt
Still about reviewer comments

2018EA000388 Answers to the first Reviewer

l. 43
Rephrased: see the new version of the manuscript

l. 57
We agree with the reviewer: the paragraph has been deleted

l. 68
Rephrased: see the new version of the manuscript

l. 68-69
IPDA informally discussed about proposing recommended data formats for archive,
but no conclusions are available at the moment.
As an IPDA official position is missing we cannot argue about FITS in that context.

l. 83-84
Rosetta data at PSA have been cited.

l. 102
GDAL lacks a metadata description/dictionary for FITS format: added in the manuscript

Table 1: note b
We agree with referee: the statement is confusing. Our goal here is to show that FITS metadata
describe a number of projections used in Planetary surface data and the mapping to PDS4
vocabulary is straightforward. The statement has been deleted.

l. 144
Planetary surface applications of standard keywords have been detailed. See the manuscript.

l. 151
WCSNAME is not the same than Table 1 column 1.
WCSNAME is a human-readable short description of the coordinate system e.g. body-fixed rotating,
-ocentric -ographic, non-rotating, inertia,... ). In PDS3 it was probably COORDINATE_SYSTEM_NAME + COORDINATE_SYSTEM_TYPE

l. 152
Rephrased: see the new version of the manuscript

l. 175
Rephrased: see the new version of the manuscript

l. 177
CUNIT is not mandatory in FITS standard: when absent, degrees are assumed.
This is a valid assumption for planetary surfaces too. 
Rephrased: see the new version of the manuscript

Table 2 : header
Mercury must be there... sorry for the oversight.
TNO and KBO are Asteroids: this is now clarified in the note.